---
title: "Research on COVID-19 vaccination"
author:
- familyname: Anukarnsakulchularp
  othernames: Krisanat
  address: Monash University
  email: kanu0003@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Vatsal
  othernames: Kumar
  address: Monash University
  email: kvat0001@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Zhou
  othernames: Xiyun 
  address: Monash University
  email: xzho0054@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Actuarial Studies
- familyname: Hu
  othernames: Xinyu
  address: Monash University
  email: xhu103@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Applied Economics and Econometrics  
department: Subject ETC5513\newline Group Name\newline Quad Squad
organization: Monash University
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r , echo = FALSE, message = FALSE, warning = FALSE}
# Libraries

library(tidyverse)
library(readr)
library(kableExtra)
library(bookdown)

```


\clearpage
# **Data set introduction** 

The data set for this report was obtained from the World Bank database, and it contains the following information:

  \begin{itemize}
    \item 
    \item 
    \item 
    \item 
    \item 
    \item 
  \end{itemize}


\clearpage

# **Research questions** :




\clearpage


# **Exploratory data analysis** 

```{r}
library(tidyverse)
```

```{r}
dat <- read_csv(here::here("Data/owid-covid-data.csv")) %>%
  select(continent, date, total_cases, new_cases, total_deaths, new_deaths, stringency_index) %>%
  na.omit()

dat
```


\section*{Effects of government policies on the spread of COVID-19 worldwide}

\subsection*{Stringency Index}

Before the availability of the vaccination, the most commons measures against the COVID-19 is the use of restrictions implemented by the government. The Oxford Covid-19 Government Response Tracker (hale_angrist_goldszmidt_kira_petherick_phillips_webster_cameron-blake_hallas_majumdar_et al._2021) calculates the Stringency Index, which is a composite measure of nine of the response metrics. The higher index score indicates a stricter response (i.e. 100 = strictest response).

```{r stringency-index, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 3, fig.cap = "The Stringency Index overtime for each Continent"}
dat %>%
  select(continent, date, stringency_index) %>%
  group_by(month = lubridate::floor_date(date, "month"), continent) %>%
  summarise(mean = mean(stringency_index),
            .groups = 'drop') %>%
  ggplot(aes(x = month, y = mean, colour = continent)) +
  geom_line() +
  theme_bw() +
  labs(x = "Time", y = "Stringency Index", color = "Continent", title = "The stingency index overtime")
```

From Figure \@ref(fig:stringency-index) we can see that each continent has a response to COVID-19 at a different time. Asia continent is the first continent to have the restriction, this is because the COVID-19 was first identified in China which is in Asia. Then the WHO (@world health organization) declared COVID-19 as a public health emergency of international concern on 30 January 2020, that is when we can see that other continents start to apply the restriction. Then on 11 March 2020 WHO assessed that COVID-19 could be characterized as a pandemic that is the month before we see the spike in the stringency index. As mentioned in the section above, there is three outbreak in the plot, every time there is a COVID-19 outbreak the stringency index also increases.

\subsection*(Effect of restriction on the total number of COVID cases)

```{r relationship-tot-index, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 3, fig.cap = "Relationship between total cases(log transformed) and stringency index"}
dat %>%
  select(continent, total_cases, stringency_index) %>%
  ggplot(aes(x = stringency_index, y = log(total_cases))) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~continent, scales = "free") +
  labs(x = "Stringency Index", y = "Total Cases (log transformation)", 
       title = "Relationship between Total Cases and Stringency index")
```

We used log transformation on the number of cases to transform the skewed data. In Figure \@ref (fig: relationship-tot-index), we can see that there is a non-linear relationship between total cases and stringency index. The total case number keeps climbing even after implementing the restriction. Then it dropped as the restriction get stricter. This is because most of the time the government will start with the lower level of restriction but this won’t have much effect on reducing the number of COVID-19 cases since the transmission of the COVID will occur before the lockdown started. Then, as the restriction increases and the longer people are staying at home, the number of total cases decreases. Therefore, we see the number of total cases climbing at first before the number goes down.

**CONCLUSIONS**

