---
title: "Research on COVID-19 vaccination"
author:
- familyname: Anukarnsakulchularp
  othernames: Krisanat
  address: Monash University
  email: kanu0003@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Vatsal
  othernames: Kumar
  address: Monash University
  email: kvat0001@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Zhou
  othernames: Xiyun 
  address: Monash University
  email: xzho0054@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Actuarial Studies
- familyname: Hu
  othernames: Xinyu
  address: Monash University
  email: xhu103@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Applied Economics and Econometrics  
department: Subject ETC5513\newline Group Name\newline Quad Squad
organization: Monash University
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      fig.align = "center",
                      fig.width = 6, 
                      fig.height = 3.5)
```



```{r , echo = FALSE, message = FALSE, warning = FALSE}
# Libraries

library(tidyverse)
library(readr)
library(kableExtra)
library(bookdown)
library(dplyr)
library(knitr)
library(ggplot2)

```


```{r echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
# Read data
dat <- read.csv(file = "Data/COVID_dat_after_sanitize.csv") 

# NA summary
summary(is.na(dat))
``` 

\clearpage
# **Data set introduction** 

The data set for this report was obtained from the World Bank database, and it contains the following information:

  \begin{itemize}
    \item 
    \item 
    \item 
    \item 
    \item 
    \item 
  \end{itemize}


\clearpage

 **Background** :






\clearpage
**Research Question 1**
\section*{Effects of government policies on the spread of COVID- 19 worldwide}







\clearpage
**Research Question 3**

\section*{ How do positive cases change relate to vaccination?}


\subsection*{top 5 countries with highest daily new cases}


```{r tab1, echo = FALSE, message = FALSE, warning = FALSE,fig.cap= "top 10 countries with highest daily cases"}
sum_dat1<-dat %>% 
  group_by(location)%>%
  mutate(new_people_fully_vaccinated = people_fully_vaccinated- lag(people_fully_vaccinated))%>%
  summarise(mean_daily_cases = round(mean(new_cases),digit = 0),
         mean_new_vax = round(mean(new_people_fully_vaccinated, na.rm = TRUE,),digits = 0))%>%
  arrange(desc(mean_daily_cases))
  kable(head(sum_dat1,10), "latex", caption = "top 10 countries with highest daily cases", booktabs = T) %>%
  kable_styling(latex_options = c("striped", "hold_position"))

```



```{r  echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
top5 <- dat%>%
  filter(location %in% c("United States", "India", "France", "Germany", "Brazil"))%>%
  select(!X)
```

\subsection*{Trend of trend of new cases vs fully vaccinated}
```{r fig2, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide' ,fig.cap = "trend of new cases vs fully vaccinated"}
p1<-ggplot(data = top5, aes(x = base::as.Date(date), y = new_cases, colour = location ))+
  geom_line() +
  theme_bw() +
  labs(y = "New Cases", x="Date")

p2<-ggplot(data = top5, aes(x = base::as.Date(date), y = people_fully_vaccinated, colour = location))+
  geom_line() +
  theme_bw() +
  labs(y = "People Fully Vaccinated", x="Date")
       

                
gridExtra::grid.arrange(p1, p2, nrow=2)    
```

\clearpage

\subsection*{Trend of trend of new cases vs fully vaccinated}

```{r fig3, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, results = 'hide' ,error = FALSE, fig.cap = "correlation between new cases vs fully vaccinated"}
ggplot(data = top5, aes(x=people_fully_vaccinated, y = new_cases))+
  geom_point()+
  geom_smooth()+
  labs(x = "people fully vaccinated" , y = "new daily cases")+
  facet_wrap(~location, scales = "free")

```

We sorted out countries by the order of mean new daily cases, referring \@ref(tab: tab1), the table showcases top 10 countries with highest mean value of daily cases, our research onward will based on the top 5 countries, US, Brazil, India, Germany, France. From table \@ref(tab:tab1), United States has the highest daily cases and India did the best in terms of vaccination.

Line chart \@ref(fig:fig2) showcases the trend of new cases and fully vaccinated people from 2021 to latest data, and from the line chart, there is no corresponding effect between them. @lipsitch2020understanding mentioned that the effectiveness of vaccination varies within different condition of health, however, both direct and indirect protection reduce virus symptoms generally. 

Plot chart \@ref(fig:fig3) was generate to explore relationship between vaccination and new cases, with vaccination on x axis, daily cases in y axis, the graph did not show significant correlation between them, smooth line was added to overview general movement. the graph further prove there is no expected higher vaccination with lower infections in the top 5 countries. @chen2021prediction discussed the how mutation reduce the effectiveness of vaccination in terms of infection protection, and the research team mentioned we need to develop vaccine to deal with predicted mutation.

\clearpage
**Research Question 3**
\section*{How do ICU rates from COVID-19 relates to population of vaccination}


\clearpage
**Research Question 4**
\section*{ How do death rates from COVID-19 differ between people who are vaccinated and those who are not}



\clearpage

**CONCLUSIONS**

The vaccination seems to not affect the spread of the virus due to mutation.

