---
title: "Research on COVID-19 vaccination"
author:
- familyname: Anukarnsakulchularp
  othernames: Krisanat
  address: Monash University
  email: kanu0003@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Vatsal
  othernames: Kumar
  address: Monash University
  email: kvat0001@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Zhou
  othernames: Xiyun 
  address: Monash University
  email: xzho0054@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Actuarial Studies
- familyname: Hu
  othernames: Xinyu
  address: Monash University
  email: xhu103@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Applied Economics and Econometrics  
department: Subject ETC5513\newline Group Name\newline Quad Squad
organization: Monash University
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r , echo = FALSE, message = FALSE, warning = FALSE}
# Libraries

library(tidyverse)
library(readr)
library(kableExtra)
library(bookdown)
library(knitr)
```

```{r message=FALSE,warning=FALSE, echo=FALSE}
data <- read.csv(here::here("Data/owid-covid-data.csv"))

```


\clearpage
# **Data set introduction** 

The data set for this report was obtained from the World Bank database, and it contains the following information:

  \begin{itemize}
    \item 
    \item 
    \item 
    \item 
    \item 
    \item 
  \end{itemize}


\clearpage

# **Research questions** :




\clearpage


# **Exploratory data analysis** 

\section*{}






\clearpage


\section*{}

**Research Question 1**



\clearpage





**Research Question 2**







\clearpage

**Research Question 3**

```{r table, echo=FALSE, message=FALSE, warning=FALSE}
data2 <- data %>% 
  filter(location %in% c("United States","Brazil","India","Russia","Mexico")) %>% 
  select(location,continent,date,people_fully_vaccinated,new_deaths,new_vaccinations,new_cases,total_cases,population)


table1 <- data2 %>%  
  group_by(location) %>% 
  summarise(average_vaccinations_per_day = round(mean(new_vaccinations, na.rm = TRUE )), average_deaths_per_day = round(mean(new_deaths, na.rm = TRUE) ) ) %>% 
  arrange(desc(average_deaths_per_day))

 kable(table1, caption = "Average daily vaccinations and deaths for the countries with highest death rates.") %>% 
  kable_classic(full_width = F, html_font = "Cambria") %>% 
    kable_styling(latex_options = c("striped", "hold_position"))

```

Table \@ref(tab:table) shows the average daily vaccinations and deaths for the countries with highest death rates. We can see that **United States** has the highest deaths per day of *1264* while **India** has the highest average vaccinations per day of approx *4057187*.


```{r fig.cap="Relationship between new deaths and vaccination status " , echo=FALSE, message=FALSE,warning=FALSE}
ggplot(data2, aes(x= (people_fully_vaccinated), 
                  y = new_deaths, color = location)) + geom_point() +
  facet_wrap(~location, scales = "free" )+
  theme_dark() +
  theme(axis.text.x = element_text(angle = 45,hjust = 1)) +
  xlab("Number of people fully vaccinated")+
  ylab("New Deaths")

```





\section*{}




\section*{}



**CONCLUSIONS**

